\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[swedish,english]{babel}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{varioref}
\usepackage{amssymb}
\usepackage{csquotes}
\usepackage{verbatim}
\usepackage{authblk}
\let\institute\affil

\usepackage[natbib,style=numeric-comp,maxbibnames=99]{biblatex}
\addbibresource{spuriouslab.bib}

\usepackage[lncs,crypto]{bibsp}
\usepackage{cleveref}

%\printanswers
\pagestyle{plain}

\title{Lab: May the (brute) force be with you, a lab to reflect on forcing 
  solutions}
\author{%
  Daniel Bosk
}
\institute{%
  Department of Information and Communication Systems\\
  Mid Sweden University, SE-851\,70 Sundsvall
}
\date{14th April 2015}

\begin{document}
\maketitle
\begin{abstract}
  \input{abstract.tex}
\end{abstract}


\section{Introduction}

The idea of this assignment is to introduce the concept of brute forcing 
security mechanisms.
The mechanism in question here is a simple monoalphabetic cryptographic 
algorithm.
It also serves to help you reflect on deniability: how can you be sure that 
your solution is the correct one?

\subsection{Aims}

The aim of this assignment is to examine that you are:
\begin{itemize}
  \input{aims.tex}
\end{itemize}


\section{Theory}

If you do not have probability theory and statistics fresh in memory you are 
recommended to revise that.
% XXX replace Swedish text with English counterpart
The text \citetitle{kthsannolikhet} by \citet{kthsannolikhet} (in Swedish) 
treats this subject, you are recommended to read Sect.~1--4.

If you have previously taken (or are currently taking) a course on 
cryptography, the material from that course covering classical cryptography is 
enough.
% XXX replace Swedish text with English counterpart
Otherwise you are recommended to read 
\citetitle{Bosk2013itn}~\cite{Bosk2013itn} (in Swedish) or Chap.~1 in 
\citetitle{Stinson2006cta} by \citet{Stinson2006cta}.

Finally, you should read about spurious keys and unicity distance.
The recommended literature is Chap.~2 in~\cite{Stinson2006cta}.


\section{Assignment}

The first part of the assignment is to break a monoalphabetic cipher.
The intercepted text is the following:
%\begin{quote}
%  TSVCFMSFQ OÅ CFMMB LVQT TLB UBQB UÄK EÖQSQPHMB NFC OQPHQBNNFQJMH PDG 
%  NBSFNBSJL PDG LQXOSPHQBEJ.
%\end{quote}
\begin{verbatim}
  TSVCFMSFQOÅCFMMBLVQTTLBUBQBUÄKEÖQSQPHMBNFCOQPHQBNNFQJMHPDG
  NBSFNBSJLPDGLQXOSPHQBEJ
\end{verbatim}
Find the corresponding plaintext of this ciphertext.
When you have found a plaintext and the key, think about how certain you can be 
that this is indeed the correct key (and thus correct plaintext).

The second part of the assignment is about spurious 
keys~\cite[Chap.~2]{Stinson2006cta} and deniability.
By spurious keys we mean a set of \(n\) keys \(k_1, k_2, \dotsc, k_n\) which 
all decrypt a ciphertext \(c\) to meaningful plaintexts \(m_1, m_2, \ldots, 
  m_n\).
Your job is to construct such a ciphertext with \(n = 2\), i.e.\ with two 
spurious keys \(k_1\) and \(k_2\), for the cryptosystem used in the first part 
of the assignment.
The texts should be as long as possible (the longest meaningful plaintexts will 
receive an award), and they do not have to be both in Swedish or English --- 
one plaintext in English and one in Swedish is fine.

Finding a spurious key algorithmically is possible, in this case, by generating 
plaintext using \(n\)-grams.
However, using pen and paper is probably the most straightforward way, and 
probably the fastest for this short assignment.


\section{Examination}

You must submit your solutions to the assignment in a report (PDF-format) in 
the course platform.
The report must contain the following:
\begin{enumerate}
  \item The plaintext corresponding to the cryptotext given above with an 
    explanation of what makes you sure this is the correct plaintext.

%    \begin{solution}
%      % XXX construct a ciphertext which can decrypt into two plaintexts
%      % XXX change the plain- and ciphertexts to English
%      The key is: \enquote{badcfehgjilknmporqtsvuwyxzåäö}.
%      The plaintext is \enquote{studenter på denna kurs ska vara väl förtrogna 
%      med programmering och matematik och kryptografi}.
%      They should motivate why this is a probably-correct decryption.
%    \end{solution}

  \item One ciphertext \(c\), two keys \(k_1\) and \(k_2\) and the 
    corresponding plaintexts \(m_1\) and \(m_2\), such that \(\Enc[k_1]{m_1} 
      = c = \Enc[k_2]{m_2}\).
    (It is not acceptable that either string contains incomprehensible text.)
    Explain your method for creating \(c, k_1, k_2, m_1. m_2\).
    Also discuss why we want to have spurious keys and how the length of the 
    message affects the spurious keys.

%    \begin{solution}
%      Verify that \( \Enc[k_1]{m_1} = c = \Enc[k_2]{m_2}\).
%    \end{solution}
\end{enumerate}


\printbibliography{}
\end{document}
