\chapter{Smart home energy management}\label{SmartEnergyStatement}
\chapterprecis{%
  This chapter is from DD2520 Applied Crypto, spring 2026, at KTH\@.
}

Many grid operators in Sweden are introducing a \emph{power tariff}
(in Swedish known as \emph{\foreignlanguage{swedish}{effekttariff}},
sometimes anglicized as \emph{effect tariff}).
Swedish regulation originally required all grid operators to apply such
tariffs from 1 January 2027, implementing EU requirements on
cost-reflective network tariffs; that mandate was revoked in 2026,
pending a revised tariff model, but cost reflectivity remains an EU
requirement and many grid operators apply power tariffs voluntarily.
With a power tariff, the electricity bill is based on the price of
electricity at the time of use, but also on the maximum power used during
the billing period.
This is done to encourage users to distribute their electricity use over time,
so that they don't run the oven, charge their electric car, run the dishwasher
and laundry machine all at the same time.

To help households distribute their electricity use, we want smart products to
coordinate when they consume power.
For instance, the car charger and the heating system can automatically lower
their consumption when they detect that the oven, vacuum cleaner, dishwasher
or laundry machine are running.
The dishwasher can delay its start until power consumption is lower or the
price is cheaper.

\ltnote{%
  We want the students to learn
  \LOapply*
  \LOanalyse*

  The object of learning is to identify which security, privacy, and
  functionality requirements arise in the smart-home setting before we discuss
  any protocol details.
  We use a more open form of pQBL~\parencite{pQBL} by starting with questions.
  This lets the reader explore the problem from their current understanding
  before we tell them more explicitly, which should reinforce the later
  explanation~\parencite{Szekely1950}.
  It also helps the reader to discern relevant
  aspects~\parencite{NecessaryConditionsOfLearning}.

  What varies is the perspective on the same household scenario: first
  benefits and drawbacks, then possible mitigations.
  What remains invariant is the underlying smart-energy setting.
  The critical aspects are confidentiality, integrity, authentication,
  authorization, privacy, and data minimization.
  The intention of
  \cref{EnergyBenefitsDrawbacks,EnergyAlleviateDrawbacks} is therefore to let
  the reader discover the properties that the coming protocol design must
  capture.
}
\begin{exercise}\label{EnergyBenefitsDrawbacks}
  What are the benefits and drawbacks of having smart devices coordinate
  their electricity use?
\end{exercise}

Typical benefits include lower electricity bills by avoiding peak-power
charges,
reduced strain on the electricity grid, and automatic comfort optimization
without manual intervention.

Drawbacks include the risk of exposing private habits through power-usage
data (when you cook, when you're home, when you sleep), the risk of
unauthorized control of devices (an attacker turning off your heating in
winter), and dependence on the security of wireless communication.

\begin{exercise}\label{EnergyAlleviateDrawbacks}
  What can we do to alleviate the drawbacks while enjoying the benefits?
\end{exercise}

We can use cryptography to protect the communication between devices:
encryption for confidentiality, authentication to prevent unauthorized
commands, and integrity protection to detect tampering.
We must also carefully consider what data the electricity provider needs and
minimize what we share.

\begin{exercise}
  Think about how you would design a system using cryptographic primitives to
  ensure the security and privacy of the smart energy system.
\end{exercise}

\ltnote{%
  The object of learning is that secure protocol design is an iterative process
  of refining a simple design by addressing one critical aspect at a time.
  We therefore begin with the simplest plausible protocol and then vary one
  design aspect per section while keeping as much else invariant as possible.
  This introduces the contrast necessary to bring each aspect into
  focus \parencite[as dictated by][]{NecessaryConditionsOfLearning}.
  The critical aspects are replay protection, direction binding, key
  separation, algorithm choice, rekeying, pairing, privacy, and trust
  assumptions (including insiders).
  The goal is not that the first protocol is good enough, but that its
  limitations make the later refinements intelligible.
  (One could also think of this as a sequence of games, as used in
  cryptography \parencite{Shoup2004Sequences}.
  Arguably, the reason why the sequences of games method by
  \citeauthor{Shoup2004Sequences} is good for humans is explained by
  \citeauthor{NecessaryConditionsOfLearning}.)
}

In this chapter, we will iteratively develop a cryptographic solution,
refining it step by step.
\chapterminitoc


\section{A first solution}\label{EnergyFirstSolution}

We have the following entities from \cref{SmartEnergyStatement}:
\begin{itemize}
  \item the power meter~\(M\), which measures the household's current power
    consumption and reports the maximum to the electricity provider,
  \item the controller~\(H\) (home hub), which coordinates electricity use
    among the devices,
  \item the smart devices~\(D_1, D_2, \ldots, D_n\), such as dishwashers,
    car chargers, heating systems and ovens,
  \item the electricity provider~\(E\), who bills the household based on
    consumption and peak power.
    (For simplicity, we let \(E\) represent both the electricity retailer
    and the grid operator.)
\end{itemize}

\begin{exercise}
  What is the goal of the system, in terms of these entities?
\end{exercise}

The goal is that the controller~\(H\) receives power-usage data from the
meter~\(M\) and from each device~\(D_i\), and sends commands to the
devices to adjust their consumption.
The provider~\(E\) receives billing data from the meter~\(M\).
All communication is wireless (radio), which means anyone nearby can
eavesdrop or inject messages.

\begin{exercise}
  Design a cryptographic system that ensures the confidentiality, integrity
  and authenticity of the communication between these entities.
\end{exercise}

The simplest design uses a shared symmetric key between the controller~\(H\)
and each device~\(D_i\).
The controller also shares a key with the meter~\(M\).

\subsection{Keys and communication}

We have a shared symmetric key~\(\KHD\) between the controller~\(H\) and
each device~\(D_i\), and a key~\(\KHM\) between the controller and the
meter~\(M\).

\begin{exercise}
  How should these keys be stored and managed?
\end{exercise}

Each device stores its key in a secure element (\eg a small \ac{TPM} or
secure microcontroller).
The controller stores all keys and must be physically protected---it is
typically installed in a locked utility cupboard or similar.

The keys can be pre-provisioned during manufacturing: each device comes with
a unique key, and the controller learns the key during a pairing procedure
(\eg by scanning a QR code on the device or by entering a code printed on
the device's label).

\subsection{A simple command protocol}

As a concrete instance, imagine that the controller tells the car charger to
\enquote{suspend charging until 14:00}.
Our first idea is to protect that command with a shared secret between the
controller and the charger.

We derive an encryption key~\(\KEHD\) and a MAC key~\(\KMHD\) from the
shared key~\(\KHD\), using a key-derivation function (such as
HKDF~\autocite{HKDF}):
\begin{align*}
  \KEHD &= \KDF[\text{\enquote{enc}}, \KHD], \\
  \KMHD &= \KDF[\text{\enquote{mac}}, \KHD].
\end{align*}

When the controller~\(H\) wants to send a command to device~\(D_i\)
(\eg \enquote{suspend until 14:00}), it encrypts and MACs the message:
\begin{align*}
  c &= \Enc[\KEHD, m], \\
  t &= \MAC[\KMHD, c].
\end{align*}

The device verifies the MAC before decrypting.
Conceptually, the tag~\(t\) authenticates the whole ciphertext~\(c\).
In practice, a MAC algorithm processes the data internally---often
incrementally and sometimes via a hash- or compression-based structure---but
the security claim is still that the tag is bound to the entire message.
Similarly, when a device reports its power consumption to the controller, it
encrypts and MACs using the same keys.

See \cref{fig:EnergyFirstProtocol} for the protocol.

\begin{figure}
  \begin{sidecaption}[Command protocol between \(H\) and \(D_i\)]
    {%
      The controller~\(H\) sends a command to device~\(D_i\).
      The message is encrypted for confidentiality and MACed for integrity
      and authentication.
      The device verifies the MAC, acknowledges authenticated receipt, executes
      the command, and finally reports the execution result.
    }[fig:EnergyFirstProtocol]
  \flushright
  \begin{msc}
    [/msc/action width=0.4\linewidth]
    {Command from \(H\) to \(D_i\)}
      \declinst{ctrl}{Controller \(H\)}{\(\KHD\)}
      \declinst{dev}{Device \(D_i\)}{\(\KHD\)}
      \mscaction{\begin{align*}
          c &= \Enc[\KEHD, m],\\
          t &= \MAC[\KMHD, c]
      \end{align*}}{ctrl}
      \nextlevel[5]
      \mess{\(c, t\)}{ctrl}{dev}
      \nextlevel
      \mscaction{\begin{align*}
          t &\stackrel ?= \MAC[\KMHD, c] \\
          m &= \Dec[\KEHD, c]
      \end{align*}}{dev}
      \nextlevel[5]
      \mscaction{If \(t\) invalid, abort; otherwise: \begin{align*}
          m_{\text{ack}} &= \text{\enquote{ack: received }} \concat m \\
          c_{\text{ack}} &= \Enc[\KEHD, m_{\text{ack}}] \\
          t_{\text{ack}} &= \MAC[\KMHD, c_{\text{ack}}]
      \end{align*}}{dev}
      \nextlevel[7]
      \mess{\(c_{\text{ack}}, t_{\text{ack}}\)}{dev}{ctrl}
      \nextlevel
      \mscaction{Execute command in \(m\).}{dev}
      \nextlevel
      \mscaction{\begin{align*}
          t_{\text{ack}} &\stackrel ?= \MAC[\KMHD, c_{\text{ack}}] \\
          m_{\text{ack}} &= \Dec[\KEHD, c_{\text{ack}}]
      \end{align*}}{ctrl}
      \nextlevel
      \mscaction{\begin{align*}
          m_{\text{report}} &= \text{\enquote{report: execution result}} \\
          c_{\text{report}} &= \Enc[\KEHD, m_{\text{report}}] \\
          t_{\text{report}} &= \MAC[\KMHD, c_{\text{report}}]
      \end{align*}}{dev}
      \nextlevel[7]
      \mess{\(c_{\text{report}}, t_{\text{report}}\)}{dev}{ctrl}
      \nextlevel
      \mscaction{\begin{align*}
          t_{\text{report}} &\stackrel ?= \MAC[\KMHD, c_{\text{report}}] \\
          m_{\text{report}} &= \Dec[\KEHD, c_{\text{report}}]
      \end{align*}}{ctrl}
  \end{msc}
  \end{sidecaption}
\end{figure}

\subsection{An example run}

\begin{exercise}
  Walk through an example run of the protocol.
  For instance, the controller sends a \enquote{suspend charging until 14:00}
  command to the car charger~\(D_{\text{charger}}\).
\end{exercise}

Let's have a look at that run:
\begin{enumerate}
  \item The controller~\(H\) prepares the command \(m = \text{\enquote{suspend 
    until 14:00}}\).
  \item It derives the encryption key~\(\KEHD\) and MAC key~\(\KMHD\) from
    the shared key~\(\KHD\).
  \item It encrypts the command: \(c = \Enc[\KEHD, m]\).
  \item It computes the MAC: \(t = \MAC[\KMHD, c]\), authenticating the
    whole ciphertext~\(c\) as described above.
  \item It sends the message \((c, t)\) to the device~\(D_{\text{charger}}\).
  \item The device receives \((c, t)\) and verifies the MAC:
    \(t \stackrel ?= \MAC[\KMHD, c]\).
  \item If the MAC is valid, it decrypts the command: \(m = \Dec[\KEHD, c]\).
  \item It executes the command, suspending charging until 14:00.
\end{enumerate}

If we leave it here, an attacker could drop the message and the controller 
would think that the charger will suspend charging whereas it will not.
We therefore add two further messages: an acknowledgment of authenticated
receipt, and then a separate execution report.
If the controller doesn't receive the acknowledgment, it can retry or alert the
user.
\begin{enumerate}[resume]
  \item The device constructs an acknowledgment message \(m_{\text{ack}} =
    \text{\enquote{ack: received suspend until 14:00}}\).
  \item It encrypts and MACs the acknowledgment:
    \(c_{\text{ack}} = \Enc[\KEHD, m_{\text{ack}}]\),
    \(t_{\text{ack}} = \MAC[\KMHD, c_{\text{ack}}]\).
  \item It sends \((c_{\text{ack}}, t_{\text{ack}})\) back to the
    controller.
  \item The controller verifies the acknowledgment and confirms that the
    command was received and authenticated by the device.
  \item After executing the command, the device sends a protected execution
    report, such as \enquote{charging suspended until 14:00}.
  \item The controller verifies that report and only then concludes that the
    command was executed successfully.
\end{enumerate}

\subsection{Consequences of the design}\label{EnergyConsequencesFirst}

\begin{exercise}
  Evaluate the design.
  What are its consequences (advantages and disadvantages) that we must
  consider?
\end{exercise}

The design has several consequences:
\begin{enumerate}
  \item Simple and efficient: symmetric cryptography is fast, which suits
    resource-constrained devices like light bulbs and appliances.
  \item Key management is manual: the user must pair each device individually
    with the controller.
  \item No replay protection: an attacker who records a valid
    \enquote{suspend} command can replay it later to disrupt the household.
  \item We have not yet said what concrete encryption mode realizes
    \(\Enc\), nor what it requires from IVs or nonces.
  \item If any device is compromised, the attacker learns the key~\(\KHD\)
    and can impersonate the controller to that device.
    But other devices are not affected (different keys).
  \item We must choose messages \(m\) carefully to prevent ciphertexts (and
    thus messages) from being reused for different purposes.
\end{enumerate}


\section{Protecting against replay attacks}\label{EnergyReplayProtection}

\ltnote{%
  The object of learning is that confidentiality and authenticity are not the
  same as freshness.
  We create contrast with the previous design where messages could be replayed.
  What varies is whether the protocol contains explicit freshness information.
  What remains invariant is the same shared-key setting and the same protected
  payload.
  The critical aspect is that encryption and authentication alone do not
  prevent replay---a separate mechanism (sequence numbers, timestamps, or
  nonces) is needed.
  A second, related critical aspect is that replay protection is not
  timeliness: a delayed message is not a replay, so expiry times are a
  separate mechanism.
  This contrast was inspired by a student solution from spring 2026.
}

In the first design, an attacker can record a valid encrypted and MACed
command and replay it later.
For example, replaying a \enquote{suspend charging} command to the car
charger every evening could prevent the car from ever charging.

\begin{exercise}
  How can we prevent replay attacks in this protocol?
\end{exercise}

To see the problem more concretely, suppose the controller once sent a valid
encrypted command meaning \enquote{suspend charging tonight} and the device
accepted it.
If an attacker records the pair \((c, t)\), then the attacker can later send
that same pair again.
The replayed message is still confidential and still has a valid tag, so the
device has no way to tell, from encryption and MAC verification alone, that
this is an old command rather than a fresh one.

\subsection{Sequence numbers}

We add a monotonically increasing sequence number~\(s\) to each message.
The sender increments~\(s\) for each message sent.
The receiver keeps track of the highest sequence number received and rejects
any message with a sequence number that is not strictly greater.

The sequence number is sent in the clear so that the receiver can check
freshness before any cryptographic processing, and the MAC covers both the
sequence number and the ciphertext:
\begin{align*}
  c &= \Enc[\KEHD, m], \\
  t &= \MAC[\KMHD, s \concat c].
\end{align*}
The transmitted message is thus \((s, c, t)\).
The receiver first checks that \(s\) is strictly greater than the last
accepted sequence number, then verifies the tag~\(t\), and only then
decrypts~\(c\).
The stored sequence number is updated only after successful verification,
so that forged messages with large sequence numbers cannot desynchronize
the channel.

Alternatively, we can use timestamps as freshness values.
The receiver then stores the timestamp from the last accepted packet and
rejects any packet whose timestamp is not strictly greater.
Used this way, timestamps act just like counters, so clock synchronization is
not required.
But then a plain sequence counter is usually simpler and more efficient.

\begin{exercise}
  What happens if the controller or a device loses power and resets its
  sequence counter?
\end{exercise}

If a device resets, its sequence counter returns to zero.
The controller would then reject its messages (sequence number too low).
To handle this, the sequence counter should be stored in non-volatile memory.
Alternatively, the device and controller can run a re-synchronization
protocol after a reset, but this must be authenticated to prevent an
attacker from forcing a counter reset.

\subsection{Delayed messages and expiry}

\begin{exercise}
  Sequence numbers ensure that an old message cannot be \emph{replayed}.
  But what if a message is merely \emph{delayed}?
  Suppose the command \enquote{suspend charging until 14:00} is sent at
  13:00 but, due to retransmissions, arrives only at 15:00.
  Does the sequence number protect against this?
\end{exercise}

No.
The delayed command has never been accepted before, so as long as no later
message has been accepted in the meantime, its sequence number is still
strictly greater than the last accepted one, and the device will accept it.
Freshness in the sense of \enquote{not previously accepted} is not the same
as timeliness.
For commands whose meaning depends on time, the sender should therefore
include an expiry time in the message, covered by the MAC, and the receiver
should reject commands that arrive after their expiry.

\subsection{Consequences of the design}

\begin{enumerate}
  \item Replay attacks are prevented: old messages are rejected.
  \item Sequence counters must be stored in non-volatile memory to survive
    power loss.
  \item Out-of-order delivery is not supported: if messages arrive out of
    order, they will be rejected.
    This is acceptable for command-and-control protocols but may need a
    sliding window for high-throughput telemetry.
  \item Delayed messages are still accepted: time-dependent commands need
    an expiry time in addition to a sequence number.
  \item The remaining issues from \cref{EnergyConsequencesFirst} still
    apply.
\end{enumerate}


\section{Binding messages to a direction}
\label{EnergyDirectionBinding}

\ltnote{%
  We now focus on another critical aspect of protocol messages: their
  direction.
  The object of learning is that a message is not only a payload with
  freshness, but also belongs to a role and a direction.
  We keep the actors, shared-secret setting, and replay-protection mechanism
  invariant while varying whether direction is represented explicitly.
  This creates contrast between a protocol where direction is only implicit
  in the surrounding prose and one where direction is cryptographically bound
  to each message.
  Following try-first pedagogy, we first ask the students to reason about
  whether two counters---one in each direction---are enough on their own.
  The generalization will come later: direction can be introduced in several
  ways, such as in the nonce, in authenticated metadata, or by deriving
  separate keys.
}

In \cref{EnergyReplayProtection}, we introduced sequence numbers to stop
replays.
Suppose now that the controller and the device each maintain one counter for
the messages they send.
Then both directions may use the values \(0, 1, 2, \ldots\).
If direction is not represented explicitly, the protocol leaves it unclear
whether the same sequence number under the same key is meant for the same or
for different purposes.

\begin{exercise}
  If both the controller and the device send a message with sequence
  number~\(0\), using the same keys, what tells the receiver which direction
  the message belongs to?
  Is the sequence number alone enough?
\end{exercise}

The sequence number alone is not enough.
Freshness tells us whether a message is new, but not whether it is a command,
an acknowledgment, or a report, nor whether it was meant for the
\(H \rightarrow D_i\) direction or the \(D_i \rightarrow H\) direction.
Direction is part of the contract of the protocol and should therefore be
bound cryptographically, not only inferred from context.

\subsection{Direction as part of the nonce}

One simple way to do this is to include the direction in the nonce.
Let \(\mathit{dir} \in \{H \rightarrow D_i, D_i \rightarrow H\}\) denote the
direction and let \(s\) be the sequence number used in that direction.
We define the nonce as
\[
  n = \mathit{dir} \concat s.
\]
Then encryption and authentication become
\begin{align*}
  c &= \Enc[\KEHD, n, m], \\
  t &= \MAC[\KMHD, n \concat c].
\end{align*}
Note that \(\Enc\) now takes the nonce~\(n\) as an explicit input alongside
the key and the message.
What the encryption mode requires from this input is the topic of
\cref{EnergyEncryptionMode}.

Now the two directions no longer reuse the same nonce under the same key,
even if both sides use the same counter value.
The receiver also verifies the nonce together with the ciphertext, so a
message valid in one direction is not automatically valid in the other.

\begin{exercise}
  Why do we authenticate the nonce~\(n\) together with the ciphertext,
  instead of only authenticating~\(c\)?
\end{exercise}

If the nonce were not authenticated, an attacker could tamper with the
direction field or the sequence number in transit.
The ciphertext might then still be delivered to the wrong cryptographic
context.
By authenticating the nonce, we bind both freshness and direction to the
message.

\subsection{Consequences of the design}

\begin{enumerate}
  \item The protocol now states explicitly which direction a message belongs
    to.
  \item Reusing the same counter value in opposite directions is safe, since
    the nonce values are still different.
  \item The protocol depends on a precise encoding of the direction field:
    both sides must agree on exactly how \(H \rightarrow D_i\) and
    \(D_i \rightarrow H\) are represented.
  \item Direction is now bound in the message format.
    In the next step, we will see that it can also be bound in the keys.
\end{enumerate}


\section{Separating keys by direction}
\label{EnergyDirectionalKeys}

\ltnote{%
  The previous section established the critical aspect that direction must be
  represented explicitly.
  We now keep that aspect invariant while varying how we realize it.
  This is a generalization pattern: the same security property can be
  achieved in more than one way.
  Here we move the directional information from the nonce into the key
  derivation, which gives students contrast between two design realizations
  of the same principle.
}

In \cref{EnergyDirectionBinding}, we encoded direction directly in the
nonce.
That works, but it is not the only possibility.
Another common way to bind direction is to derive different keys for the two
directions.

\begin{exercise}
  Can we express the direction of communication in the key material instead
  of in the nonce?
  What would that buy us?
\end{exercise}

Yes.
Starting from the long-term shared key~\(\KHD\), we derive separate keys for
encryption and MACing in each direction:
\begin{align*}
  \KEHD^{\rightarrow} &= \KDF[\text{\enquote{enc to device}}, \KHD], \\
  \KEHD^{\leftarrow} &= \KDF[\text{\enquote{enc to controller}}, \KHD], \\
  \KMHD^{\rightarrow} &= \KDF[\text{\enquote{mac to device}}, \KHD], \\
  \KMHD^{\leftarrow} &= \KDF[\text{\enquote{mac to controller}}, \KHD].
\end{align*}
Here \(\rightarrow\) means traffic from \(H\) to \(D_i\), and
\(\leftarrow\) means traffic from \(D_i\) to \(H\).

Now each direction uses its own sequence number and its own keys:
\begin{align*}
  c &= \Enc[\KEHD^{\rightarrow}, s^{\rightarrow}, m], \\
  t &= \MAC[\KMHD^{\rightarrow}, s^{\rightarrow} \concat c]
\end{align*}
for commands from \(H\) to \(D_i\), and
\begin{align*}
  c &= \Enc[\KEHD^{\leftarrow}, s^{\leftarrow}, m], \\
  t &= \MAC[\KMHD^{\leftarrow}, s^{\leftarrow} \concat c]
\end{align*}
for acknowledgments and reports from \(D_i\) to \(H\).

Because the keys are different, the same counter value can safely appear in
both directions without causing nonce reuse under one and the same key.
We have made direction part of the cryptographic context by domain-separating
the derived keys.

\begin{exercise}
  Compare the two ways of representing direction:
  as part of the nonce or as part of the key derivation.
  What are the advantages and disadvantages of each?
\end{exercise}

Putting direction in the nonce keeps the key schedule simpler but makes the
message format more explicit.
Putting direction in the key derivation gives strong domain separation and
lets each direction use a plain counter as nonce, but it increases the number
of keys that must be derived, stored, and implemented correctly.

\subsection{Consequences of the design}

\begin{enumerate}
  \item The two directions are cryptographically separated even if they use
    the same counter values.
  \item Implementations become less likely to confuse commands with
    acknowledgments or reports, since they use different keys.
  \item The protocol now uses more key material and requires more careful key
    management inside the controller and each device.
  \item We still have not said what concrete encryption mode
    \(\Enc\) should be, nor what it requires from IVs, nonces, and
    randomness.
\end{enumerate}


\section{Choosing an encryption mode}
\label{EnergyEncryptionMode}

\ltnote{%
  The object of learning is that choosing \(\Enc\) means choosing both a mode
  of operation and the corresponding nonce/IV discipline.
  We keep the protocol structure invariant and vary only the concrete
  realization of \(\Enc\).
  The critical aspect is that different secure block-cipher modes require
  different things from the extra input: some need unpredictable IVs, others
  only unique nonces.
  Following try-first pedagogy, we first let students compare plausible mode
  choices before telling them which requirements matter in this setting.
  What varies is the mode and its input requirements; what remains invariant is
  the same command-and-control setting.
  A further critical aspect, inspired by a student solution from spring
  2026, is that the nonce space is finite: counters wrap, so rekeying is
  part of nonce discipline.
}

So far we have written \(\Enc\) as if it were a single black box.
In practice, we must choose a concrete encryption mode.
That choice matters because different modes leak different information and
place different demands on IVs, nonces, and randomness.

\begin{exercise}
  Suppose two commands begin with the same header, such as
  \enquote{command: suspend ...}.
  If we encrypt them block by block in the most direct way, what information
  might an eavesdropper still learn?
\end{exercise}

If we use \ac{ECB}, equal plaintext blocks become equal ciphertext blocks.
Then repeated structure in commands, acknowledgments, or reports is visible to
an eavesdropper even though the content is encrypted.
For a command-and-control protocol, this may reveal message types or repeated
operations.
So \ac{ECB} is not a suitable choice here.

\subsection{Different modes, different requirements}

\begin{exercise}
  If we replace \(\Enc\) by a better block-cipher mode, what extra value must
  usually accompany the plaintext?
  Must that value be random, unique, or both?
\end{exercise}

Different modes have different requirements:
\begin{description}
  \item[\ac{CBC}] Requires a fresh, unpredictable IV for each encryption.
    Reusing or predicting the IV can reveal information about related
    plaintexts.
  \item[\ac{CTR}] Requires a nonce that is unique for each encryption under a
    given key.
    The nonce need not be unpredictable, but it must never repeat.
  \item[\ac{AEAD}] Modes such as \ac{GCM} combine encryption and integrity
    protection, but they still rely on a nonce discipline.
    In particular, nonce reuse under one key is dangerous.
\end{description}

This difference is important in our smart-home setting.
Many devices are cheap, resource-constrained, and may not have a strong random
number generator available every time they send a message.
That makes modes requiring fresh unpredictable IVs harder to justify.
By contrast, our protocol already maintains sequence numbers in each
direction, so we can reuse those counters as nonces for modes that only need
uniqueness.

\subsection{Where should the randomness come from?}

\begin{exercise}
  Is it reasonable to expect a simple smart plug or thermostat to generate a
  high-quality random IV every time it sends a message?
  If not, how can we avoid relying on that capability?
\end{exercise}

It is risky to rely on strong runtime randomness in low-end devices.
Some devices have weak entropy sources, especially just after boot.
If the security of the mode depends on a fresh unpredictable IV each time,
then implementation mistakes or poor randomness can break confidentiality.

Our current design therefore points toward a mode that tolerates deterministic
nonce generation, provided the nonce is unique under each key.
With per-direction keys from \cref{EnergyDirectionalKeys} and per-direction
counters, the protocol already provides such a uniqueness source.
This is another reason why counter-based modes fit this setting well.

Long-term key generation is a different issue.
The shared secret~\(\KHD\) should not be generated by an ad hoc weak random
process inside a cheap device during normal operation.
Instead, long-term keys should come either from a proper provisioning process
during manufacturing or from a later pairing procedure that is designed to
establish keys securely.

\subsection{Switching to authenticated encryption}

\begin{exercise}
  If we already use encryption and a MAC, why might we still prefer an
  authenticated-encryption mode instead of combining two separate
  primitives ourselves?
\end{exercise}

Using separate encryption and MAC primitives can be secure, but only if they
are combined correctly.
There are many ways to get this wrong: authenticating the wrong fields,
forgetting to authenticate the nonce, or using a less suitable order of
operations.
An \ac{AEAD} mode~\autocite{Rogaway2002aead} packages confidentiality and
integrity into one well-analyzed primitive and therefore reduces the number
of design choices that an implementer must get right.

Two common \ac{AEAD} modes are
AES-\ac{GCM}~\autocite{McGrewViega2004gcm,SP80038D} and
ChaCha20-Poly1305~\autocite{RFC8439}.
With authenticated encryption, a protected command from \(H\) to
\(D_i\) has the form
\begin{align*}
  c &= \AEADEnc[\KEHD^{\rightarrow}, n, m, a],
\end{align*}
where \(n\) is the nonce and \(a\) is associated data that is
authenticated but not encrypted.

\begin{exercise}
  What information in our protocol should be authenticated even if we do not
  want to encrypt it?
\end{exercise}

The associated data should include the context that the receiver must trust in
order to interpret the message correctly, such as the sequence number, sender
and receiver identities, direction, and protocol version.
In our setting, this fits naturally with the earlier sections on replay
protection and direction binding.

This does not remove the need for nonce discipline.
\ac{AEAD} still requires that the nonce be managed correctly.
For modes such as AES-\ac{GCM}, nonce reuse under one key is particularly
dangerous.
Our per-direction counters are therefore still useful: they provide a natural
nonce source once direction and key separation have already been introduced.

\subsection{Rekeying before the counter wraps}

\begin{exercise}
  Our nonces are counters of some fixed width.
  What happens when a counter reaches its maximum value?
  What must we do before that happens?
\end{exercise}

When the counter wraps around, the next nonce repeats an earlier one.
Nonce reuse under the same key is exactly what we must avoid; for modes
such as AES-\ac{GCM} it can be catastrophic for both confidentiality and
authenticity.
We must therefore rotate the keys before the counter wraps: for example,
derive fresh keys from the long-term key~\(\KHD\) with an epoch number as
an additional \(\KDF\) input, and reset the counters when the epoch
changes.
In practice, implementations rotate keys after a fixed time or number of
messages, with the counter width chosen so that wrapping never happens in
normal operation.
Recall from \cref{EnergyReplayProtection} that the counters must also
survive reboots, otherwise a device may unknowingly reuse nonces after a
power loss.

\subsection{Consequences of the design}

\begin{enumerate}
  \item \ac{ECB} is unsuitable because it leaks repeated block patterns.
  \item Some secure modes require unpredictable IVs, while others require only
    nonce uniqueness.
  \item In this setting, per-direction counters are an attractive nonce source
    because they avoid dependence on strong runtime randomness.
  \item \ac{AEAD} reduces composition mistakes, but it still depends on
    correct nonce management.
  \item Counters wrap eventually: keys must be rotated before the nonce
    space is exhausted.
\end{enumerate}


\section{Should we just use TLS?}
\label{EnergyTLS}

\ltnote{%
  We now compare our purpose-built protected message protocol with a standard
  secure-channel protocol.
  The critical aspect is that transport assumptions matter: TLS gives strong
  security properties, but it assumes a reliable stream transport, whereas
  DTLS is designed for datagrams.
  Following try-first pedagogy, we first ask whether a secure channel can be
  reused for sparse traffic, and then make the contrast between TLS and DTLS
  explicit.
  We don't expect students to have detailed knowledge of how TCP keeps
  connections alive, nor to know of DTLS.
  We do expect them to reason about the problem for the low-energy devices;
  rather than proposing DTLS, they will likely assume that TLS over TCP can
  keep a channel open indefinitely without keep-alive packets---giving the
  same result as using DTLS.
}

At this point, a natural question is whether we should stop designing our own
message protection and simply use Transport Layer Security,
TLS~1.3~\autocite{RFC8446}, or Datagram Transport Layer Security,
DTLS~1.3~\autocite{RFC9147}.

\begin{exercise}
  If a smart device sends only one short message per hour, would using TLS
  mean performing a full handshake for every message?
  If not, what practical difficulty remains?
\end{exercise}

TLS does not require a handshake for every message.
If the connection remains open, the endpoints can establish a TLS session once
and then send protected application data whenever needed.
However, TLS assumes a reliable, in-order byte stream, typically TCP.
That means the protected channel exists only while the underlying connection
exists.

For sparse smart-home traffic, this creates practical problems.
If the device keeps the TCP/TLS connection open for long idle periods, network
equipment may time it out, forcing keepalives or reconnects.
If the device instead reconnects whenever it has data, it pays the cost of
setting up TCP and then running a new TLS handshake, perhaps with session
resumption but still with non-trivial state and latency.
Session resumption helps, because it avoids the cost of a full fresh
authentication and key exchange every time, but it does not remove the need to
re-establish transport state after the old connection has disappeared.

\begin{exercise}
  If our traffic is naturally packet-oriented rather than stream-oriented,
  which is the better fit: TLS or DTLS?
  Why?
\end{exercise}

DTLS~1.3~\autocite{RFC9147} is the closer match for message-oriented
traffic.
It adapts the TLS~1.3 design to datagram transports and therefore does not
require a long-lived TCP-style stream.
That makes DTLS conceptually attractive for command-and-report traffic where
messages are naturally separate packets.
However, DTLS still has a handshake, state, retransmission logic, and
anti-replay mechanisms.
In particular, the receiver must still track packet numbers and reject old or
duplicate packets, so replay protection has not disappeared---it has merely
been packaged inside the secure-channel design.
So it is not free; it is simply better aligned with datagram communication
than TLS over TCP.

\begin{exercise}
  If standard protocols such as TLS and DTLS already exist, when would a
  purpose-built protocol still be justified?
\end{exercise}

A custom protocol may still be justified when the devices are extremely
constrained, the traffic pattern is simple and narrowly scoped, and the system
has requirements that do not fit comfortably into a stream-oriented or general
datagram secure channel.
But the price is that we must design, implement, and analyze replay
protection, direction binding, nonce management, algorithm choices, and error
handling ourselves.
In other words, the custom design is attractive only if the gain in simplicity
or efficiency is large enough to compensate for the extra design risk.

In our setting, the comparison becomes a design trade-off:
\begin{description}
  \item[TLS~1.3 (RFC~8446)] Attractive for richer devices and for links that
    already have a stable TCP connection.
    It is especially natural for the meter's communication with the provider,
    where an Internet-style client/server model already exists.
  \item[DTLS~1.3 (RFC~9147)] More natural for sparse packet-based traffic
    inside the home, because it does not impose a byte-stream abstraction.
    It still carries handshake, retransmission, anti-replay, and
    state-management complexity.
  \item[Purpose-built protocol] Can be lighter and more tailored to the
    devices, but is justified only if the constraints are strong enough to
    outweigh the safety of reusing a standard protocol.
\end{description}

\subsection{Consequences of the design}

\begin{enumerate}
  \item A standard secure-channel protocol can reduce design risk by reusing a
    well-analyzed construction.
  \item TLS is awkward for sleepy devices if traffic is sparse and idle
    connections are expensive to maintain.
  \item DTLS is a better fit than TLS for datagram-style traffic, but it still
    does not remove the need for handshake state, retransmissions,
    authentication, anti-replay tracking, and key establishment.
  \item A purpose-built protocol is easiest to justify only when the device
    constraints are severe and the communication pattern is very simple.
  \item The remaining problem is still how a new device obtains the shared
    key~\(\KHD\) in the first place.
    That is the purpose of the next section on pairing.
\end{enumerate}


\section{Device authentication and pairing}\label{EnergyPairing}

\ltnote{%
  The object of learning is that secure pairing is a bootstrapping problem:
  how do you establish trust and a shared secret when there is no prior shared
  secret?
  We create contrast with the previous assumption that keys are
  pre-provisioned.
  What varies is the way trust is established during onboarding.
  What remains invariant is the later protected communication protocol.
  The critical aspects are authentication of the other endpoint, resistance to
  impersonation, and the usability cost of the onboarding step.
}

In \cref{EnergyFirstSolution}, we assumed that each device is pre-provisioned
with a key known to the controller.
In practice, a user buys a new smart light bulb and wants to add it to their
home network.
The bulb has no prior relationship with the controller.

\begin{exercise}
  How can a new device securely join the home network and establish a shared
  key with the controller?
\end{exercise}

\subsection{Out-of-band pairing}

The standard approach is \emph{out-of-band} (OOB) pairing: using a separate,
trusted channel to bootstrap the key exchange.
Common methods:
\begin{description}
  \item[QR code] The device has a QR code (printed on its body or packaging)
    containing a pre-shared secret or a public key.
    The user scans it with the controller's app.
  \item[NFC tap] The user holds the device near the controller (or a phone
    acting as the controller's interface) to exchange pairing data via
    \ac{NFC}.
  \item[PIN entry] The device displays or is labelled with a PIN code.
    The user enters it into the controller.
\end{description}

After the OOB exchange provides a shared secret (or authenticated public
keys), the devices can run a key-agreement protocol (\eg Diffie--Hellman)
over the wireless channel, authenticated by the OOB secret, to establish
the long-term shared key~\(\KHD\) that we previously assumed to be
pre-provisioned.

If this setup step is not authenticated properly, a nearby attacker can mount
a man-in-the-middle attack during onboarding.
Consider a neighbour who is annoyed by the dishwasher running at night.
If the neighbour inserts themselves between the controller and the dishwasher
during setup, the dishwasher may appear to work normally as long as the
neighbour relays the messages.
Later, the neighbour can simply stop relaying in the evening or at night,
preventing the controller from starting the dishwasher.
The same attack works against a smart-speaker system: the speaker functions
while the neighbour relays traffic, but stops responding whenever the
neighbour chooses not to forward the messages.

To detect this kind of setup attack, we do not necessarily need mutual
authentication.
Authenticating one side is already enough: if either the controller can
authenticate the device, or the device can authenticate the controller, the
attacker cannot invisibly replace that authenticated endpoint.
This is analogous to HTTPS/TLS, where authenticating the web server is enough
to detect a man-in-the-middle attack even when the client is not authenticated
to the server.

The unauthenticated endpoint can still be impersonated outright, however:
if the device cannot authenticate itself, the attacker can pair their own
radio with the controller, while the real device never joins.
But this is not an invisible compromise either: the real device remains
unpaired and unresponsive, so the user notices that onboarding failed.
(In HTTPS, the analogous substitution is harmless for a different reason:
the unauthenticated endpoint is the user's own client, which the attacker
cannot usefully replace.)

Note carefully what authentication does and does not buy us here.
With an authenticated key exchange, the attacker cannot learn the
established key, so they cannot read or forge messages---they can only
relay ciphertexts unmodified or refuse to relay them.
Refusing to relay is then no more powerful than jamming the radio channel
(\cref{EnergyWirelessSecurity}): a blunt, detectable availability attack
rather than an invisible compromise.

\begin{exercise}
  What are the security properties of each pairing method?
  Which is most suitable for the different types of devices in our system?
\end{exercise}

\begin{description}
  \item[QR code] Works well for devices that are too simple for a display or
    keypad (\eg light bulbs).
    The code is static, so if an attacker photographs it, they can pair
    later.
    The code should be inside the packaging and treated as confidential
    until pairing.
  \item[NFC tap] Requires proximity (a few centimetres), which provides
    physical authentication.
    Good for devices the user can physically approach.
    Harder to intercept than a QR code.
  \item[PIN entry] Requires a display or label on the device and an input
    method on the controller.
    A short PIN (\eg 6 digits) resists brute force only if the pairing
    protocol limits guessing: each protocol run must admit at most one
    PIN guess (as in a password-authenticated key exchange) and failed
    pairing attempts must be rate-limited.
    Suitable for devices with displays (\eg ovens, car chargers).
\end{description}

\subsection{Certificate-based onboarding}

For a more scalable approach, especially in larger installations, we can use
certificates~\autocite{Kohnfelder1978}.
Each device is manufactured with a device certificate signed by the
manufacturer's \ac{CA}.
The controller~\(H\) has its own key pair~\((\SKH, \VKH)\), certified by a
home \ac{CA} trusted by the devices, or else its public key~\(\VKH\) is
authenticated out-of-band during onboarding.
If the controller uses a self-signed certificate, that certificate is useful
only once the device has learned to trust \(\VKH\) through such an out-of-band
step.

During pairing:
\begin{enumerate}
  \item The device presents its manufacturer certificate.
  \item The controller verifies the certificate chain.
  \item The controller presents either a certificate under the trusted
    home \ac{CA} or a public key that the device authenticates
    out-of-band.
  \item The device verifies that controller credential.
  \item They perform authenticated key exchange (using their respective
    keys) to establish the shared key~\(\KHD\).
\end{enumerate}

This approach doesn't require QR codes or PINs for each session, but it does
require explicit trust roots on both sides: the controller must trust the
manufacturer's \ac{CA}, and the device must trust the controller's
onboarding credential.
It also assumes the manufacturer has not been compromised and that device
certificates have not been cloned.

\subsection{Consequences of the design}

\begin{exercise}
  Evaluate the pairing methods.
  What are the trade-offs?
\end{exercise}

\begin{description}
  \item[Usability vs.\ security] QR codes and NFC are easy for users but
    rely on physical security; certificate-based pairing is seamless but
    requires trust in manufacturers.
  \item[Setup MITM risk] If onboarding authenticates neither side strongly
    enough, a nearby attacker can place themselves between the controller and
    the device during pairing and later disrupt availability by selectively
    stopping relayed traffic.
  \item[Scalability] Certificate-based pairing scales better to many
    devices, since no per-device manual step is needed.
  \item[Device cost] Certificates require more capable hardware (crypto
    coprocessor) in each device, increasing cost.
    Symmetric-key OOB pairing can work with very cheap hardware.
  \item[Revocation] If a device is compromised, we need a way to revoke its
    key.
    With pre-shared keys, the controller simply deletes the key.
    With certificates, a revocation mechanism is needed.
\end{description}


\section{Protecting wireless communication}\label{EnergyWirelessSecurity}

\ltnote{%
  The object of learning is that a secure message format does not by itself
  secure the wireless medium.
  We create contrast with the previous focus on message-level security
  (encryption and MACs).
  What remains invariant is the same protected protocol.
  What varies is the threat that comes from the channel itself: availability in
  the jamming case, and metadata leakage in the traffic-analysis case.
  The critical aspects are therefore separate threat dimensions rather than one
  single wireless-security property.
}

All communication in the system happens over wireless radio.
Our encrypted and authenticated messages protect against eavesdropping on
content and against message forgery.
But there are additional threats specific to wireless channels.

\begin{exercise}
  What threats exist at the wireless channel level, beyond what our
  encryption and authentication already address?
\end{exercise}

\subsection{Jamming}

An attacker can transmit noise on the radio frequency to prevent all
communication.
This is a denial-of-service attack.
If the devices cannot communicate, the controller cannot coordinate
power use, potentially causing the household to exceed its power limit
and incur higher bills.

\begin{exercise}
  How can we mitigate jamming attacks?
\end{exercise}

Jamming cannot be prevented by cryptographic means, so complete protection
is out of scope for our cryptographic design, but mitigations include:
\begin{description}
  \item[Frequency hopping] The devices agree (during pairing) on a
    pseudo-random frequency-hopping sequence.
    An attacker must jam all frequencies simultaneously, which requires
    much more power.
    This is standard in protocols like Bluetooth.
  \item[Graceful degradation] Devices should have sensible defaults when
    they cannot reach the controller.
    For example, the car charger could continue charging at a low rate
    rather than stopping entirely.
  \item[Detection and alerting] The controller can detect communication
    failures and alert the user.
\end{description}

\subsection{Traffic analysis}

Even with encryption, an attacker observing the wireless traffic can learn:
\begin{itemize}
  \item Which devices are communicating (by radio fingerprinting or
    observing communication patterns).
  \item When commands are sent (indicating changes in household activity).
  \item The volume of data, which may reveal whether devices are active.
\end{itemize}

\begin{exercise}
  How can we mitigate traffic analysis?
\end{exercise}

Mitigations include:
\begin{description}
  \item[Constant-rate traffic] Devices transmit at a fixed rate,
    sending dummy messages when there is no real data.
    This hides when actual commands are sent.
  \item[Uniform message sizes] Pad all messages to the same length
    so that message type cannot be inferred from size.
  \item[Randomized timing] Add random delays to message
    transmission to obscure patterns.
\end{description}

These measures increase power consumption and bandwidth use, so they must
be weighed against the limited resources of battery-powered devices.

\subsection{Consequences of the design}

\begin{description}
  \item[Availability] Frequency hopping and graceful degradation improve
    resilience against jamming.
  \item[Privacy] Traffic-analysis countermeasures protect activity patterns
    but consume more energy and bandwidth.
  \item[Complexity] Frequency hopping requires shared state between devices
    and controller; constant-rate traffic requires careful timing management.
  \item[Battery life] Continuous dummy traffic significantly reduces battery
    life for battery-powered devices (sensors, smart plugs).
    A reasonable trade-off may be to use constant-rate traffic only for
    sensitive devices.
\end{description}


\section{Privacy of energy data}\label{EnergyPrivacy}

\ltnote{%
  The object of learning is that privacy is not only about protecting data in
  transit, but also about limiting what data is disclosed to legitimate
  external parties.
  This creates contrast with the previous focus on securing communication
  between the controller and devices.
  What remains invariant is the local smart-home protocol.
  What varies is the granularity of the data shared with the provider.
  The critical aspect is that the electricity provider~\(E\) learning detailed
  consumption data is a different threat from a wireless eavesdropper learning
  ciphertexts and metadata.
}

The power meter~\(M\) records the household's power consumption over time.
The electricity provider~\(E\) needs this data for billing.
But detailed consumption data reveals intimate information about the
household's daily life.

\begin{exercise}
  What can an electricity provider (or an attacker who compromises the
  meter data) learn from detailed power consumption data?
\end{exercise}

From fine-grained power data, one can infer:
\begin{itemize}
  \item When residents are home and when they are away.
  \item When they wake up and go to sleep.
  \item What appliances they use (each appliance has a characteristic power
    signature).
  \item Whether they have an electric car and when they charge it.
  \item How many people live in the household (from shower and cooking
    patterns).
\end{itemize}

This is called \emph{non-intrusive load monitoring} (NILM) and is a
well-studied technique~\autocite{Hart1992nilm}.

\begin{exercise}
  How can we design the system to minimize the privacy impact while still
  enabling the provider to bill correctly?
\end{exercise}

\subsection{Data minimization}

The principle of \emph{data minimization} (central to the \ac{GDPR})
states that we should collect only the data strictly necessary for the
purpose.

For billing under effect tariffs, the provider needs:
\begin{itemize}
  \item Total energy consumed during the billing period.
  \item Maximum power observed during the billing period.
\end{itemize}

The provider does \emph{not} need minute-by-minute or second-by-second
consumption readings.
Therefore, the meter should report only aggregated values.

\subsection{Aggregation at the meter}

The meter~\(M\) can compute the aggregated billing data locally and report
only the result to the provider~\(E\):
\begin{itemize}
  \item Total consumption: \(\sum_{t} P(t) \cdot \Delta t\).
  \item Peak power: \(\max_{t} P(t)\) over the billing period.
\end{itemize}

The meter sends these values to the provider, encrypted and signed, at the
end of each billing period.
The detailed time series stays within the household.

\begin{exercise}
  The controller~\(H\) needs detailed data from the meter for coordination.
  How do we handle this without leaking data to the provider?
\end{exercise}

The meter~\(M\) shares detailed readings with the controller~\(H\) over the
local wireless network, protected by the key~\(\KHM\).
The provider only receives the aggregated billing data.
This separation is enforced by the meter's software, which must be trusted.

\subsection{Consequences of the design}

\begin{exercise}
  Evaluate the privacy design.
  What are its strengths and weaknesses?
\end{exercise}

\begin{description}
  \item[Privacy] The provider learns only aggregate billing data, not
    detailed usage patterns.
    This is a significant improvement over reporting fine-grained data.
  \item[Trust in the meter] The meter must be trusted to correctly compute
    aggregates and not leak detailed data.
    This is a single point of trust.
  \item[Provider's perspective] The provider cannot verify that the meter
    is reporting honestly.
    We address this in \cref{EnergyMeteringIntegrity}.
  \item[Local privacy] Within the household, the controller sees detailed
    data.
    If the controller is cloud-connected (\eg for remote access), this
    data could leak to the cloud provider.
\end{description}


\section{Integrity of metering data}\label{EnergyMeteringIntegrity}

\ltnote{%
  The object of learning is the tension between privacy and verifiability in
  metering.
  We create contrast with the previous section where the meter was simply
  trusted.
  What remains invariant is the privacy goal of reporting only aggregates.
  What varies is whether the provider receives cryptographic evidence for those
  aggregates.
  The critical aspect is that data minimization and provider-side confidence
  are in tension, and that signatures plus hardware trust are one way to bridge
  that tension.
}

The electricity provider~\(E\) needs to trust that the billing data from the
meter~\(M\) is correct.
A dishonest household could tamper with the meter to report lower consumption
or a lower peak power, reducing their bill.

\begin{exercise}
  How can the provider verify that the meter's reports are correct?
\end{exercise}

\subsection{Signed meter readings}

Each meter~\(M\) has a signing key pair~\((\SKM, \VKM)\), provisioned during
manufacturing.
The public key~\(\VKM\) is registered with the provider~\(E\) (possibly via
the manufacturer's certificate chain).

Let
\[
  r = (\mathit{id}, \mathit{period}, W_{\text{tot}}, P_{\text{peak}},
  \mathit{ts})
\]
be the billing report.
The meter signs that report:
\[
  \sigma = \Sign[\SKM, r].
\]

The provider verifies:
\(\Ver[\VKM, r, \sigma] \stackrel{?}{=} \text{true}\).

This ensures that the report was produced by the authentic meter and has not
been modified in transit.
Conceptually, the signature authenticates the whole report~\(r\).
In practice, modern signature schemes first hash the report and then apply the
public-key operation to that short digest.
This is also why public-key cryptography is normally used to authenticate
reports, identities, or freshly established symmetric keys, while the actual
bulk payload is protected with symmetric cryptography.

\begin{exercise}
  The signature proves the report came from the meter, but what if the
  meter's hardware has been physically tampered with?
\end{exercise}

\subsection{Tamper-evident meters}

Physical security measures complement the cryptographic ones:
\begin{description}
  \item[Tamper-evident seals] Visible indicators that the meter's enclosure
    has been opened.
  \item[Tamper-responsive hardware] The meter's secure element erases the
    signing key if physical intrusion is detected, making further signed
    reports impossible.
  \item[Secure boot] The meter verifies its own firmware integrity at
    start-up, refusing to operate if the software has been modified.
  \item[Remote attestation] The meter can prove to the provider that it is
    running genuine, unmodified firmware.
\end{description}

\subsection{The communication channel to the provider}

\begin{exercise}
  How should the meter communicate with the electricity provider?
\end{exercise}

The meter can communicate with the provider through several channels:
\begin{description}
  \item[Power-line communication (PLC)] Data transmitted over the
    existing electrical wiring.
    Common in smart metering deployments.
  \item[Cellular (4G/5G)] The meter has its own SIM card and
    communicates directly with the provider.
  \item[The household's internet connection] Cheaper but relies
    on the household's network, which the household controls.
\end{description}

Regardless of the channel, the meter authenticates to the provider and
establishes an encrypted connection (\eg TLS with mutual authentication).
The signed billing reports provide an additional layer of integrity that
holds even if the transport channel is compromised.

\subsection{Consequences of the design}

\begin{description}
  \item[Billing integrity] Signed reports provide origin authentication and
    accountability: the provider can attribute a valid report to the meter
    (assuming the meter hardware is intact).
  \item[Privacy preserved] The provider still receives only aggregate data;
    the signature doesn't require sharing more information.
  \item[Hardware cost] Tamper-evident meters with secure elements are more
    expensive than simple meters.
  \item[Key management] The provider must maintain a database of all meter
    public keys and manage certificate revocation for decommissioned meters.
  \item[Single device trust] The entire billing integrity depends on the
    meter's hardware and software being trustworthy.
    This is a strong assumption, but it is standard practice in metering.
\end{description}


\section{Firmware updates and device lifecycle}\label{EnergyFirmware}

\ltnote{%
  The object of learning is that even a well-designed protocol depends on a
  trustworthy implementation.
  We create contrast with the protocol-level security we have focused on so
  far.
  What remains invariant is the protocol design itself.
  What varies is whether the device software executing that design can be
  replaced or modified by an attacker.
  The critical aspect is therefore the trust chain from firmware signing to
  secure boot.
}

Smart devices, the controller, and the meter all run software (firmware) that
implements our protocols.
This firmware will need updates to fix bugs, patch security vulnerabilities,
and add features.

\begin{exercise}
  What threats are associated with firmware updates?
  How can an attacker exploit the update mechanism?
\end{exercise}

An attacker who can install malicious firmware on a device can:
\begin{itemize}
  \item Extract cryptographic keys from the device.
  \item Disable security checks (MAC verification, encryption).
  \item Cause the device to malfunction (\eg the heating system running
    at maximum in summer).
  \item Use the device as a foothold to attack other devices on the
    network.
\end{itemize}

\subsection{Signed firmware updates}

Each manufacturer has a firmware-signing key pair
\((\SKManuf, \VKManuf)\).
The device stores the manufacturer's public key~\(\VKManuf\) in its secure
element.

Before installing a firmware update, the device verifies:
\[
  \Ver[\VKManuf, f, \sigma_{\text{fw}}]
  \stackrel{?}{=} \text{true}.
\]

Here \(f\) denotes the whole firmware image.
As with meter signatures, the verification claim is about the entire object
being authenticated, even though the underlying signature algorithm will in
practice hash that object internally before applying the public-key operation.

If verification fails, the update is rejected.

\subsection{Secure boot}

\begin{exercise}
  Even if we only install signed firmware, what happens if the device's
  storage is modified directly (by physical access)?
\end{exercise}

\emph{Secure boot} provides the answer: at every power-on, the device
verifies the integrity of its own firmware before executing it.
The chain of trust starts from an immutable boot ROM that verifies the
bootloader, which in turn verifies the main firmware.

If any stage fails verification, the device refuses to boot normally and
enters a recovery mode that only accepts signed firmware from the
manufacturer.

\subsection{Revoking compromised devices}

\begin{exercise}
  If a device is compromised (its keys extracted by an attacker), how
  should the system respond?
\end{exercise}

The controller should maintain a list of authorized devices.
When a device is reported compromised:
\begin{enumerate}
  \item The controller removes the device's key~\(\KHD\) from its
    database, refusing further communication.
  \item If certificate-based pairing is used, the device's certificate is
    added to a revocation list.
  \item The user is notified to replace or reset the device.
  \item If the compromised device's key was used for any shared state,
    affected keys should be rotated.
\end{enumerate}

\subsection{Limiting what insiders can do}\label{EnergyInsiders}

\ltnote{%
  The object of learning is that cryptographic authentication establishes
  who sent a message, not that its contents are true or safe.
  We keep the protocol invariant and vary which party is compromised: first
  a device lying in its own telemetry, then the controller itself.
  The critical aspects are the difference between channel integrity and
  data integrity, and between authentication and authorization.
  This contrast was inspired by a student solution from spring 2026.
}

Revocation handles a compromised device once we know about it.
But how do we notice, and how do we limit the damage in the meantime?

\begin{exercise}
  A compromised device cannot forge messages from other devices, since each
  device has its own key.
  But it can lie in its own reports.
  What can a compromised device achieve by reporting false power readings,
  and how can the controller detect this?
\end{exercise}

By reporting false readings, a compromised device can manipulate the
coordination: for example, reporting a huge consumption may cause the
controller to throttle all the other devices.
The MAC only proves that the report comes from the device---it says nothing
about whether the reading is true.
Cryptography protects the channel, not the honesty of the sensor.

The controller can, however, cross-check: the meter~\(M\) independently
measures the household's total consumption.
The devices' reported readings can never legitimately sum to more than the
meter's total, since the household also has loads that are not smart
devices.
If the sum exceeds the total, some device is over-reporting, and the
controller can alert the user.
(A device that under-reports hides in the unmetered remainder, so this
check is one-sided.
It also assumes consumption only: in a household with local generation,
such as solar panels, devices can legitimately consume more than the
meter measures from the grid.)

\begin{exercise}
  What if the controller itself is compromised?
  It holds all keys and can send valid commands to every device.
\end{exercise}

The controller is the most privileged party, so its compromise is the most
damaging.
We cannot prevent a compromised controller from sending valid commands, but
we can limit what those commands achieve: each device should enforce local
safety limits and reject commands that violate them, even when the command
is cryptographically valid.
For example, the heating system should refuse to turn off completely in
winter, and the car charger should never exceed its electrical rating.
A valid MAC proves who sent the command, not that the command is safe:
authentication is not authorization.

\subsection{Consequences of the design}

\begin{description}
  \item[Software integrity] Signed firmware and secure boot prevent
    unauthorized software from running on devices.
  \item[Manufacturer trust] Devices must trust their manufacturer's
    signing key.
    If a manufacturer's key is compromised, all their devices are
    vulnerable.
  \item[Update delivery] Firmware updates must reach all devices reliably,
    including battery-powered devices that may be offline for extended
    periods.
  \item[Backward compatibility] Updated firmware must maintain protocol
    compatibility with devices that haven't been updated yet.
  \item[Device lifecycle] Eventually, manufacturers stop supporting old
    devices.
    There should be a mechanism to continue operating (with a clear
    security status) or to decommission them gracefully.
  \item[Defense in depth] Local safety limits in each device bound the
    damage of a compromised controller; cross-checks against the meter
    bound the damage of a lying device.
\end{description}


\section{Legal and usability considerations}\label{EnergyLegalUsability}

\ltnote{%
  The object of learning is that a technically plausible cryptographic design
  must still be evaluated against legal and usability constraints.
  This creates contrast with the purely technical focus of previous
  sections.
  What remains invariant is the same technical system.
  What varies is the evaluation perspective: compliance and everyday use.
  The critical aspects are that a system can fail in practice even if its core
  cryptography is sound, because it may violate legal requirements or be too
  hard for ordinary households to operate correctly.
}

\begin{exercise}
  What legal requirements apply to a smart energy management system that
  processes household energy data?
\end{exercise}

\subsection{GDPR and energy data}

Under the \ac{GDPR}, energy consumption data is personal data because it
can reveal information about the household's inhabitants.
Key obligations:
\begin{description}
  \item[Purpose limitation] Data may only be processed for specified
    purposes (billing, grid management).
  \item[Data minimization] Collect only what is strictly necessary, as we
    designed in \cref{EnergyPrivacy}.
  \item[Storage limitation] Detailed data should not be retained longer
    than needed.
  \item[Data subject rights] Households can request access to their data,
    correction, or deletion.
  \item[Security] Appropriate technical measures (our cryptographic
    protocols) must protect the data.
\end{description}

\subsection{EU smart meter regulations}

The EU sets requirements for smart metering in the Electricity
Directive~\autocite[Articles~19--24]{EUElectricityDirective}:
\begin{itemize}
  \item Meters must be able to be read remotely by the provider.
  \item Consumers must have access to their own consumption data.
  \item Data must be provided in a standardized format.
  \item Consumers can choose to share their data with third parties (\eg
    energy-management services).
\end{itemize}

Our design must ensure that the consumer can share data selectively,
without giving the third party access to more data than necessary.

\subsection{Usability for non-technical households}

\begin{exercise}
  How do we make the system usable for households where nobody has
  technical expertise?
\end{exercise}

Key usability principles:
\begin{description}
  \item[Automatic coordination] Devices should coordinate without
    requiring the user to understand the underlying protocols.
    The system should work out of the box with sensible defaults.
  \item[Simple pairing] Adding a new device should require at most scanning
    a QR code or pressing a button.
    The user should not have to configure keys, algorithms, or
    security parameters.
  \item[Transparent operation] The user should see the system's effect
    (\enquote{you saved X kr this month by shifting dishwasher runs to
    off-peak hours}) without needing to understand how.
  \item[Graceful failure] If a device cannot communicate, it should
    continue operating at reasonable defaults rather than failing silently
    or requiring manual intervention.
  \item[Clear error messages] If something goes wrong, messages should be
    actionable: \enquote{the dishwasher lost contact with the controller,
    check that the controller is powered on} rather than
    \enquote{AEAD authentication failure on frame 4712}.
\end{description}

\subsection{Consequences of the design}

\begin{description}
  \item[Compliance] The system must be designed with the \ac{GDPR} and EU
    metering regulations in mind from the start, not as an afterthought.
  \item[Usability--security tension] Making the system easy to use may
    require relaxing some security measures (\eg allowing pairing with
    just a button press, which provides weaker authentication than a
    QR code with a long secret).
  \item[Third-party data sharing] Providing APIs for third-party services
    introduces new attack surfaces and privacy risks.
    Consent management and access-token scoping are essential.
  \item[Accessibility] The system must be usable by people with
    disabilities, which may affect the choice of pairing methods
    (not everyone can scan a QR code).
\end{description}


\section{Future work and open problems}\label{EnergyFutureWork}

\ltnote{%
  The closing exercise is a fusion task: the student picks one open
  problem and must coordinate the critical aspects practiced one at a
  time in this chapter---threat modelling, primitive choice, key
  management, and trade-off analysis---in a new, less structured
  setting.
  The list of topics varies the application context while the iterative
  design method of the chapter remains invariant, which is intended to
  support transfer.
}

Several aspects of smart energy management systems remain active areas of
research and development:

\begin{description}
  \item[Post-quantum cryptography] The symmetric algorithms (AES,
    HMAC) used in device-to-controller communication are largely
    quantum-resistant.
    However, any Diffie--Hellman-style key
    establishment~\autocite{DiffieHellman} must be migrated to a
    post-quantum key-establishment mechanism such as
    ML-KEM~\autocite{FIPS203}, and any certificate/signature layer must
    likewise migrate to a post-quantum signature scheme such as
    ML-DSA~\autocite{FIPS204} to ensure long-term security.
  \item[Denial-of-service on the radio channel] Jamming and flooding
    attacks on the wireless channel can disrupt coordination.
    Advanced techniques such as spread-spectrum communication,
    cryptographic puzzles for rate limiting, and redundant communication
    paths (\eg combining radio with power-line communication) could
    improve resilience.
  \item[Multi-household coordination] In apartment buildings, multiple
    households share the same electrical supply.
    Coordinating across households raises privacy challenges: how can
    the building optimize collective consumption without any household
    revealing its individual usage?
    Techniques from secure multi-party computation could enable this.
  \item[Renewable energy and local trading] Households with solar panels
    may want to sell excess electricity to neighbours.
    This peer-to-peer energy trading requires authenticated and
    privacy-preserving transaction protocols.
  \item[Supply-chain security for devices] How do we ensure that smart
    devices have not been tampered with during manufacturing, transport,
    or retail?
    Hardware attestation and supply-chain verification protocols are
    relevant.
  \item[Machine-learning on encrypted data] The controller could use
    machine learning to predict consumption patterns and optimize
    scheduling.
    If this processing moves to the cloud, homomorphic encryption or
    federated learning could preserve privacy.
\end{description}

\begin{exercise}
  Choose one of the topics above and outline how you would approach
  designing a solution.
  What cryptographic primitives would you use?
  What are the main challenges?
\end{exercise}


\section*{Acknowledgements}

This chapter was written by Daniel Bosk with the assistance of Dan-Claude
van Thropic.

Thanks to the students of the spring 2026 course for their inspiring 
submissions (random order):
Luca Manzari, Amin Nouiser, Fredrik Bjälå Ursäter, Elvis Tran, David Schalin, 
Jens Jonsson, William Malmberg, Elias Vitell, Hasim Demirok, Eric Lexenberg, 
Wiktor Dobrosierdow, Love Lindgren, Josef Anhede, Silvia Lin Lü, Vanja 
Grigoriev, Linnéa Wennberg, Anna Akopyan, Mathias Magnusson, Zhongmin Hu, 
Tobias Bjurström, Ahmedhadi Bashir Ibrahim, Chloe Joyce Wabingga Raneses, 
Oliver Bojs, Jimmy Tran, Tenzin Sangpo Choedon, Ivar Boqvist, Elias 
Kristiansson, Felix Winkelmann, Fredrik Magnevill, Vilhelm Prytz, Seerat Yasir, 
Daniel Norman, Axel Axhag, Benjamin Widman, Daniel Nikolic, Jonathan Blomlöf, 
Anders Mouanga, Mohammed Faisal Ahmed Ba Rashed, Karl Walfridson, David 
Byström, Daniel Svensson, Marcus Pramborg, Naveed Rahman, Felix Bergqvist 
Widström, Theodor Björkman, Jonathan Nilsson Cullgert, Henrik Nur, Denise 
Hommik, Viktor Sandström, Zeyuan Zhao, Mulham Abo Khalaf, Adam Maeedi, Vivienne 
Curewitz, Samreen Sidhu, Lorenzo Deflorian, Tamás Ferenc Fülöp, Simone 
Brunello, Sarah Susanne Winter, Haik Hovhannesian, Shenyu Dai, Anton Holderied, 
Manuel Michael Voit, Paula Oramas Badillo, Nils Kovacic, Alexandru Carp, Aurora 
Navas Jimenez, Robert Rittner, Venetia Ioanna Papadopoulou, Ana Maria Dobrescu, 
Zoi Koukouve, Baptiste Drouet, Timothée Finck, Aleksander Kregiel, Yousra Al 
Mowahed, Sumeya Yasir Isse, Jonas Ralf Jostan, Jafar Mohammadi, Gabriel 
Alejandro Arias Goa, Only Badle, Augusto Sebastian Barthelemy Solá, Ignacio 
Searles, and Romain Foucaud.
