\mode*

% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section{Public-key cryptography}

\subsection{Key-exchange schemes}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \item It's difficult to have to exchange keys in advance.

        \pause{}

      \item What if we could securely exchange keys at a distance?
      \item If we could do it just before we use them?
    \end{itemize}
  \end{idea}

%  \pause{}
%
%  \begin{exercise}
%    \begin{itemize}
%      \item Say Alice and Bob want to communicate.
%      \item Eve wants to eavesdrop as usual.
%      \item What are the requirements of such a system?
%    \end{itemize}
%  \end{exercise}
\end{frame}

\begin{frame}
  \begin{solution}[Requirements]
    \begin{itemize}
      \item We need a problem that is easy for Alice and Bob.
      \item It should be hard for Eve.
    \end{itemize}
  \end{solution}
\end{frame}

\begin{frame}
  \begin{definition}[\Acl{DLP}, \acs{DLP}]
    \begin{itemize}
      \item Let \(\Z_p^*\) be the multiplicative group of residues modulo 
        \(p\in \N\), where \(p\) is a prime.

        \pause{}

        \begin{description}
          \item[Given] \(g, g^x\in \Z_p^*\)
          \item[Find] \(x\).
        \end{description}

      \item I.e.\ compute \(\log_{g\in \Z_p}(g^x)\).
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{definition}[\Acl{DHP}, \acs{DHP}\footfullcite{DiffieHellman}]
    \begin{description}
      \item[Given] \(g, g^x, g^y\in \Z_p^*\)
      \item[Find] \(g^{xy}\)
    \end{description}
  \end{definition}

  \pause{}
  
  \begin{definition}[\Acl{DDH} Problem, \acs{DDH}]
    \begin{description}
      \item[Given] \(g, g^x, g^y, g^z\in \Z_p^*\)
      \item[Decide] \(z \stackrel{?}{=} xy\)
    \end{description}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item If we can solve \ac{DLP}, then we can solve \ac{DHP} and \ac{DDH} 
      too.

      \pause{}

    \item Maybe \ac{DHP} and \ac{DDH} can be solved without \ac{DLP}.
    \item We don't know yet.

      \pause{}

    \item We usually assume \ac{DLP}, \ac{DHP} and \ac{DDH} are hard.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{exercise}
    \begin{itemize}
      \item \citeauthor{DiffieHellman}\footfullcite{DiffieHellman} used 
        \ac{DHP} to create a key-exchange protocol.

        \pause{}

      \item Take some time to figure out how we can use these problems to 
        achieve what we want.
    \end{itemize}
  \end{exercise}

  \begin{block}{Reminder}
    \begin{itemize}
      \item Alice and Bob want to exchange a secret key.
      \item Then they can use the key to encrypt their communications.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \begin{definition}[Diffie-Hellman key-exchange]
    \begin{itemize}
      \item Let \(g\in \Z_p^*\) (publicly known, e.g.\ RFC, standard\ dots).

        \pause{}

      \color{green!50!red}
      \item Alice generates random \(0 < x < |\Z_p^*|\).
      \item She send \(g^x\) to Bob.

        \pause{}

      \color{blue}
      \item Bob generates random \(0 < y < |\Z_p^*|\).
      \item He sends \(g^y\) to Alice.

        \pause{}

      \color{green!50!red}
      \item Alice has \(x\) and \(g, g^y\).
      \color{blue}
      \item Bob has \(g, g^x\) and \(y\).
      \color{black}
      \item They both compute \(g^{xy} = {\color{green!50!red} (g^y)^x} 
          = {\color{blue} (g^x)^y}\).

        \pause{}

      \color{red}
      \item Eve has \(g, g^x, g^y\).
      \item By \ac{DHP} she cannot compute \(g^{xy}\).
    \end{itemize}
  \end{definition}
\end{frame}

\subsection{Encryption and decryption}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \item Fine, we can use \(g^{xy}\) as a key in a cipher.
        \begin{itemize}
          \item \(\Enc[g^{xy}]{m}\), where \(\EncOp\) is a symmetric cipher.
        \end{itemize}
      \item But shouldn't we be able to include a message directly?
    \end{itemize}
  \end{idea}
\end{frame}

\begin{frame}
  \begin{definition}[ElGamal Encryption Scheme\footfullcite{ElGamal}]
    Set-up:
    \begin{itemize}
      \item Let \(g\in \Z_p^*\), randomly choose \(0 < x < |\Z_p^*|\).
      \item Alice publishes \(\Z_p^*, g, g^x\) to everyone.
    \end{itemize}
    Encryption:
    \begin{itemize}
      \item Bob chooses random \(0 < y < |\Z_p^*|\) and computes \(g^y\).
      \item Bob's message \(m\in \Z_p^*\).
      \item He sends \((g^y, m(g^{x})^y)\) to Alice.
    \end{itemize}
    Decryption:
    \begin{itemize}
      \item Alice computes \((g^y)^x\) and \(m(g^x)^y((g^{y})^x)^{-1} = m\).
    \end{itemize}
  \end{definition}
\end{frame}

\subsection{Digital signatures}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \item Sure, if Bob sends a message to Alice, he's sure she's the only one
        who can decrypt it.

        \pause{}

      \item Can't we turn this around?
        \begin{itemize}
          \item Can't Alice use the same system to ensure Bob knows the message
            came from Alice?
        \end{itemize}
    \end{itemize}
  \end{idea}

  \pause{}

  \begin{exercise}
    \begin{itemize}
      \item Look at the ElGamal encryption scheme for a bit.
      \item Try to find a way to \enquote{run it backwards}.
    \end{itemize}
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{definition}[ElGamal Signature Scheme\footfullcite{ElGamal}]
    Set-up:
    \begin{itemize}
      \item Let \(g\in \Z_p^*\) and \framebox{\(h\) be a one-way function}.
      \item Alice publishes \(\Z_p^*, g, g^x\) to everyone.
    \end{itemize}
    Signing \(m\in \Z_p^*\):
    \begin{itemize}
      \item \framebox{Alice} chooses random \(0 < y < |\Z_p^*|\) and computes 
        \(r = g^y\in \Z_p^*\).
      \item She computes \framebox{\(s = (h(m) - xr) y^{-1} \pmod{|\Z_p^*|}\).}
      \item She sends \((r, s)\) to Bob.
    \end{itemize}
    Verification:
    \begin{itemize}
      \item \framebox{Bob checks if \(g^{h(m)} \stackrel{?}{=}_{\Z_p^*} (g^x)^r 
            r^s\)}\( =_{\Z_p^*}
          (g^x)^{g^y} (g^y)^{(h(m) - xg^y)y^{-1}} =_{\Z_p^*}
          g^{xg^y + h(m) - xg^y}\)
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item It works without the hash.
      \item But then we can multiply two messages and still get a valid 
        signature.
    \end{itemize}
  \end{remark}
\end{frame}

\subsection{Homomorphic properties}

\begin{frame}
  \begin{definition}[Homomorphism]
    A \emph{homomorphism} is a map (function) that preserves structure between 
    two algebraic structures.
  \end{definition}

  \pause{}

  \begin{example}
    \begin{itemize}
      \item Let \(G_1 = (\R, \cdot)\) and \(G_2 = (\R, +)\) be groups.
      \item \(g_1, g_1^\prime\in G_1\) and \(g_2, g_2^\prime\in G_2\).

        \pause{}

      \item Consider \(\log\colon G_1\to G_2\).
        
        \pause{}

      \item \(\log(g_1\cdot g_1^\prime) = g_2 + g_2^\prime\).
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{exercise}
    The encryption (decryption) function of the ElGamal cryptosystem is 
    a homomorphism, what structure does it preserve?
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{example}[ElGamal's homomorphism]
    \begin{itemize}
      \item Messages \(m, m^\prime\), ciphertexts \((g^y, m\cdot g^{xy}), 
          (g^{y^\prime}, m^\prime\cdot g^{x y^\prime})\).

      \item Remember: private key \(x\), hence the same.

        \pause{}

      \item Create ciphertext
        \begin{align*}
          (g^y g^{y^\prime}, m\cdot g^{xy}\cdot m^\prime\cdot g^{x y^\prime})
          &= (g^{y + y^\prime}, m\cdot m^\prime\cdot g^{xy + xy^\prime}) \\
          &= (g^{y + y^\prime}, m\cdot m^\prime\cdot g^{x(y + y^\prime)}).
        \end{align*}

        \pause{}

      \item Decryption: take \(g^{y + y^\prime}\), compute \((g^{y+y^\prime})^x 
          = g^{x(y + y^\prime)}\).

      \item Decryption thus yields \(m\cdot m^\prime\).

    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item We use a hash function in the signature scheme to counter the 
        homomorphic property.

      \item \(h(m)\cdot h(m^\prime)\neq h(m\cdot m^\prime)\).

        \pause{}

      \item Without the hash function we could create a valid signature for 
        a new message \emph{without knowing the signature key!}
    \end{itemize}
  \end{remark}
\end{frame}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item There are many schemes with different homomorphic properties.
      \item There is even \emph{fully homomorphic 
          encryption}~\cite{GentryFullyHomomorphicEncryption}.
    \end{itemize}
  \end{remark}
\end{frame}


\section{More counter-intuitive things}

\subsection{Secure multi-party computation}

\begin{frame}
  \begin{example}[Yao's Millionaires' Problem]
    \begin{itemize}
      \item Two millionaires meet in the street.
      \item They want to find out who is the richer.

        \pause{}

      \item However, they don't want to reveal how many millions they each 
        have.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \uncover<1>{%
      \item We have \(n\) participants \(P_1, \ldots, P_n\).
      }
      \uncover<2>{%
      \item Each person has a \emph{secret} input value \(x_j\) for \(1\leq 
          j\leq n\).
      }
      \uncover<3>{%
      \item But they desperately want to know \(y = f(x_1, \ldots, x_n)\).
      }
    \end{itemize}
  \end{idea}
%\end{frame}
%\begin{frame}
  \begin{figure}
    \begin{tikzpicture}[->,auto,node distance=1.5cm,
      thick,input node/.style={circle,draw},
      func node/.style={rectangle,draw}]

      \uncover<1,2,4>{%
      \node[input node] (1) {$P_1$};
      \node[input node] (2) [right of=1] {\ldots};
      \node[input node] (3) [right of=2] {$P_n$};
      }

      \uncover<3,4>{%
      \node[func node] (4) [below of=2] {$f$};

      \node (5) [below of=4] {$f(x_1, \ldots, x_n) = y$};
      }

      \uncover<4>{%
      \path[every node/.style={font=\small}]
      (1) edge node [right] {$x_1$} (4)
      (2) edge node {} (4)
      (3) edge node [right] {$x_n$} (4)
      (4) edge node {} (5);
      }
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \begin{example}[Trivial solution]
    \begin{itemize}
      \item The \(n\) participants \(P_1, \dotsc, P_n\) agree on \iac{TTP}.
      \item Each participant give their secret to the \ac{TTP}.
      \item The \ac{TTP} trusted third-party performs the computation.
      \item Every participant receives the result from the \ac{TTP}.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{definition}[\Acl{MPC}, \acs{MPC}]
    \begin{itemize}
      \item \(n\) participants \(P_1, \ldots, P_n\).
      \item \(n\) secret inputs \(x_1, \ldots, x_n\).

        \pause{}

      \item A protocol \(\pi\) is executed by the participants.
      \item At the end of the protocol each participant learns \(y = f(x_1, 
          \ldots, x_n)\).

        \pause{}

      \item The participants executing \(\pi\) should be \emph{equivalent} to 
        giving \(x_1, \ldots, x_n\) to \iac{TTP} \(T\) who computes \(f(x_1, 
          \ldots, x_n) = y\) and returns \(y\) to each participant.
    \end{itemize}
  \end{definition}

  \pause{}

  \begin{remark}
    Each participant \(P_i\) learns no more about \(x_j\) (\(i\neq j\))
    than what is revealed by \(y\).
  \end{remark}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item In general this problem is solved.
    \item We can construct protocols for arbitrary functions \(f\).

      \pause{}

    \item Efficiency varies though.
    \item However, there are practically feasible protocols.

      \pause{}

    \item Sometimes we can use homomorphisms.
    \item But we can construct rather complex functions too.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{example}[Sugar beet auctions\footfullcite{MPCgoesLive}]
    \begin{itemize}
      \item Several thousand farmers produce sugar beets.
      \item These are sold to the monopoly Danisco, the sugar producer.

        \pause{}

      \item Contracts are allocated via a nation-wide exchange, a double 
        auction.

      \item A double auction contains multiple sellers and multiple buyers.

      \item The purpose is to find the \emph{market clearing price}.

    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}[Sugar beet auctions, continued]
    \begin{itemize}
      \item Each buyer places a bid specifying how much he is willing to buy 
        \emph{at each potential price}.

      \item Each seller says how much they are willing to sell at each given 
        price.

        \pause{}

      \item The auctioneer computes the total supply and demand for each price.

      \item We want to find where supply equals demand.

        \pause{}

      \item When done, anyone who specified non-zero for this price may trade 
        at this price.
    \end{itemize}
  \end{example}
\end{frame}

\subsection{Zero-knowledge proofs of knowledge}

\begin{frame}
  \begin{example}
    \begin{itemize}
      \item Alice must prove her identity to Eve.

      \item Eve has Alice's public key, and knows it belongs to Alice.

        \pause{}
        
      \item Alice wants to prove she is the owner of the private key belonging 
        to the public key that Eve has.

        \pause{}

      \item Eve asks Alice to sign the message \(m\), if the signature verifies
        under the public key Eve believes Alice.
        
    \end{itemize}
  \end{example}

  \pause{}

  \begin{alertblock}{Gaaahh!}
    \begin{itemize}
      \item Now Eve can show this message (chosen by Eve) with Alice's 
        signature on it!
      \item What if Eve's chosen message was \enquote{I give all my money to 
          Eve}?
    \end{itemize}
  \end{alertblock}
\end{frame}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \item Alice wants to prove that she knows the discrete logarithm \(x\) of 
        a value \(g^x\).

      \item She will do this without revealing \(x\) to Eve.
    \end{itemize}
  \end{idea}
\end{frame}

\begin{frame}
  \begin{definition}[Schnorr's protocol\footfullcite{Schnorr}]
    \begin{itemize}
      \item Prover wants to prove knowledge of \(x\) for \(g^x = y\).

        \pause{}

      \item Prover commits to randomness \(r\), by sending \(t = g^r\).

        \pause{}

      \item Verifier replies with randomly chosen challenge \(c\).

        \pause{}

      \item After receiving \(c\), prover replies with \(s = r + cx\).

        \pause{}

      \item Verifier accepts if \(g^s = g^{r + cx} = g^r (g^x)^c = t y^c\).
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{proof}[Proof outline]
    \begin{itemize}
      \item We need to prove \emph{completeness}: for all (most) statements the 
        verifier will accept.

        \pause{}

      \item We need to prove \emph{soundness}: for all (most) false statements 
        the verifier will reject.

        \pause{}

      \item We need to prove that it is zero-knowledge.
    \end{itemize}
  \end{proof}
\end{frame}

\begin{frame}
  \begin{block}{Zero-knowledge}
    \begin{itemize}
      \item Transcript for protocol: \((t, c, s)\).

      \item Probability for transcript occurring: \(\frac{1}{|R|}\cdot 
          \frac{1}{\deg g}\).

        \pause{}

      \item Simulate protocol: randomly choose \(c\), randomly choose \(s\), 
        compute \(t\) by \(g^s y^c\).

        \pause{}

      \item We see that we get the same probability distribution.
      \item Thus the simulated transcripts are indistinguishable from the real 
        ones.
    \end{itemize}
  \end{block}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[allowframebreaks]
  \printbibliography{}
\end{frame}

